The Categorical Structure of Profunctor Optics

### 2017-11-01 paper.tex
Tiny changes

\section{Indexed Optics}
\todo{
This possibly looks like
\begin{align*}
\int^{M \in \M} \C(S, M(IA)) \times \C(M A', S')
\end{align*}
with another monoidal action $I$. The $I$ would need to distribute over $M$ in some way. ``Coindexed'' optics seem to have 
\begin{align*}
\int^{M \in \M} \C(S, J(MA)) \times \C(M A', S')
\end{align*}
which is why they pass information `in the other direction' but still transform covariantly.
}
\section{Merely Well-Behaved Optics}
\todo{
The lenses that the UPenn crew look at are less restricted. They obey the $\fput\fget$ and $\fget\fput$ laws but not the $\fput\fput$ law. 
}

@@

\begin{example}
A minimal example is the following:
\begin{align*}
\fget : \{A,B,C\} &\to \{X, Y\} \\
A &\mapsto X \\
B &\mapsto X \\
C &\mapsto Y 
\end{align*}
\begin{align*}
\fput : \{A,B,C\} \times \{X, Y\} &\to \{A,B,C\} \\
(A,X) &\mapsto A \\
(B,X) &\mapsto B \\
(C,X) &\mapsto B \\
(A,Y) &\mapsto C \\
(B,Y) &\mapsto C \\
(C,Y) &\mapsto C
\end{align*}
\todo{So here the $\fput\fget$ composite is
\begin{align*}
(A,X) &\mapsto (A,X) \\
(B,X) &\mapsto (B,X) \\
(C,X) &\mapsto (B,X) \\
(A,Y) &\mapsto (C,Y) \\
(B,Y) &\mapsto (C,Y) \\
(C,Y) &\mapsto (C,Y)
\end{align*}
}
\end{example}

\todo{The condition on $\langle l, r \rangle$ seems to be that $rl = \id_S$ and $\pi_2lr = \pi_2$. I can't see how to generalise this second condition to optics other than lenses.}

@@

Past attempts \cite{ReflectionsOnMonadicLenses} to define effectful lenses have wrapped the result of one or both of $\fget$ and $\fput$ with a monad. This approach is a little ad-hoc; it is not obvious what the laws ought to be and there is no clear generalisation to other varieties of optic. The definition of optic given in this paper suggests we instead set $\C$ to be the Kleisli category for some monad.

>>>>

\section{Indexed Optics}
\todo{
This possibly looks like
\begin{align*}
\int^{M \in \M} \C(S, M(IA)) \times \C(M A', S')
\end{align*}
with another monoidal action $I$. The $I$ would need to distribute over $M$ in some way. ``Coindexed'' optics seem to have 
\begin{align*}
\int^{M \in \M} \C(S, J(MA)) \times \C(M A', S')
\end{align*}
which is why they pass information `in the other direction' but still transform covariantly.
}

\section{Merely Well-Behaved Optics}
\todo{
The lenses that the UPenn crew look at are less restricted. They obey the $\fput\fget$ and $\fget\fput$ laws but not the $\fput\fput$ law. 
}

@@

\begin{example}
A minimal example of a prelens that obeys $\fget\fput$ and $\fput\fget$ but not $\fput\fput$ is the following. Let $p : \{A,B,C\} \hto \{X, Y\}$ be the prelens with
\begin{align*}
\fget : \{A,B,C\} &\to \{X, Y\} \\
A &\mapsto X \\
B &\mapsto X \\
C &\mapsto Y 
\end{align*}
\begin{align*}
\fput : \{A,B,C\} \times \{X, Y\} &\to \{A,B,C\} \\
(A,X) &\mapsto A \\
(B,X) &\mapsto B \\
(C,X) &\mapsto B \\
(A,Y) &\mapsto C \\
(B,Y) &\mapsto C \\
(C,Y) &\mapsto C
\end{align*}
\end{example}

Prelenses that satisfy the first two laws correspond to pairs $\langle l, r \rangle$ such that $rl = \id_S$ and $\pi_2lr = \pi_2$. \todo{prove?} This condition seems unavoidably tied to the product structure on $\C$; it is not clear whether this weakened notion of lens can be generalised to other optics.

@@

Many attempts \cite{ReflectionsOnMonadicLenses} \todo{more} to define effectful lenses have wrapped the result of one or both of $\fget$ and $\fput$ with a monad. This approach is a little ad-hoc; it is not obvious what the laws ought to be and there is no clear generalisation to other varieties of optic. The definition of optic given in this paper suggests we instead set $\C$ to be the Kleisli category for some monad.
